---
title: "JC_Stats_rerun"
author: "Ben"
date: "5/10/2021"
output: html_document
---

```{r packages, include = F}
library(tidyverse)
library(car)
library(ggpubr)
library(kableExtra)
library(corrplot)
library(broom)
```


The goal of this document is to re-run data organization and analyses completely and transparently, independent of all the goofy functions I've written, so that our analysis process can be easily reviewed and verified.

I'm going to begin with the importation of raw data from FAO, and generate the change-over-time metrics independently for each figure and analysis.

As we discussed in our last meeting, the primary points I am trying to examine/test are:

1. Land sparing trends from 1970s - 1990s were not present in the last 20 years

2. Over the last 20 years, countries with increased nitrogen inputs were less likely to have reduced the area in staple crop production.

3. 95-’15 all countries have increased exports of staple crops, with the biggest increases in exports positively correlated with increases in staple crop land areas

4. High and very high HDI countries have shown the greatest increase in land use, yield, exports, and nitrogen use.

5. Globally there has been a reduction in total staple crop land 

## Data Importation and Cleaning
```{r FAO Data Importation}
# Importing and filtering staple crops from FAO Ag Production Data
staple_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item =="Apples"| Item == "Barley"| Item ==  "Bananas"| Item == "Cassava"| Item == "Coconuts"| Item ==  "Seed cotton"| 
                  Item ==  "Grapes"| Item ==  "Maize"| Item ==  "Millet"| Item == "Oats"| Item ==  "Onions, dry"| Item ==  "Oil palm fruit"| 
                  Item ==  "Plantains"| Item ==  "Potatoes"| Item ==  "Rice, paddy"| Item ==  "Sorghum"| 
                  Item == "Soybeans"| Item ==  "Sugar beet"| Item ==  "Sugar cane"| Item ==  "Sunflower seed"| Item ==  "Sweet potatoes"| 
                  Item ==  "Vegetables, fresh nes"| Item == "Wheat"| Item ==  "Yams") %>% 
  dplyr::filter(Element != "Production") 
names(staple_data) = gsub(pattern = "Y", replacement = "", x = names(staple_data))
staple_data_clean <- staple_data
area.tall <- pivot_longer(data = staple_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area")%>% 
  dplyr::select(Area, Item, Year,area)
area.tall$Year <- as.numeric(area.tall$Year)
head(area.tall)


```
