---
title: "JC Figures 5_20"
author: "Ben"
date: "5/19/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(car)
library(ggpubr)
library(kableExtra)
library(corrplot)
library(broom)
library(stringr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
```

```{r data importation, include = FALSE}
fao_staple_complete <- read.csv("../data/fao_updated_staple.csv", sep = ",", header = T)
mergeset <- read.csv("../data/mergeset.csv", sep = ",", header = T)
```



## 2 panel boxplot figure: 

### A. per capita agricultural land is decreasing in most countries, and B per hectare yields are increasing in most countries. This may suggest land sparing is occur… (strawman)


```{r initial data generation, include = F}
Area_Yield_Change_79_99 = fao_staple_complete %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # Filter Countries here if wanted later
    dplyr::select(Area,Year,Population,kcal.ha.avg,area.tot) %>%    # 
    dplyr::filter(!is.na(kcal.ha.avg) & !is.na(area.tot))%>%   # 
    dplyr::filter(between(Year,(1979-2),(1979+2))|between(Year,(1999-2),(1999+2))) %>%           
    dplyr::mutate(yield_percap = kcal.ha.avg, area_percap = area.tot/Population, # Removed Per-capita
                  time = if_else(Year <= (1979+2) & Year >= (1979-2),"start.5","end.5"))%>% 
    dplyr::select(Area,time,yield_percap,area_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.yield = mean(yield_percap,na.rm = T),avg.area = mean(area_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.yield,avg.area))  %>% 
    dplyr::mutate(yield_dif = avg.yield_end.5 / avg.yield_start.5,
                  area_dif = avg.area_end.5 / avg.area_start.5) %>% 
    dplyr::mutate(log.yield.dif = log(yield_dif), 
                  log.area.dif = log(area_dif)) %>% 
    dplyr::filter(!is.na(log.yield.dif) & !is.na(log.area.dif))

# Add HDI data back into dataset
Area_Yield_Change_79_99 <- left_join(Area_Yield_Change_79_99,mergeset %>% dplyr::select(1,2,4,5,7))

Area_Yield_Change_95_15 = fao_staple_complete %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # Filter Countries here if wanted later
    dplyr::select(Area,Year,Population,kcal.ha.avg,area.tot) %>%    # 
    dplyr::filter(!is.na(kcal.ha.avg) & !is.na(area.tot)) %>%   # 
    dplyr::filter(between(Year,(1995-2),(1995+2))|between(Year,(2015-2),(2015+2))) %>%           
    dplyr::mutate(yield_percap = kcal.ha.avg, area_percap = area.tot/Population, # Removed Per-capita
                  time = if_else(Year <= (1995+2) & Year >= (1995-2),"start.5","end.5"))%>% 
    dplyr::select(Area,time,yield_percap,area_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.yield = mean(yield_percap,na.rm = T),avg.area = mean(area_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.yield,avg.area))  %>% 
    dplyr::mutate(yield_dif = avg.yield_end.5 / avg.yield_start.5,
                  area_dif = avg.area_end.5 / avg.area_start.5) %>% 
    dplyr::mutate(log.yield.dif = log(yield_dif), 
                  log.area.dif = log(area_dif)) %>% 
    dplyr::filter(!is.na(log.yield.dif) & !is.na(log.area.dif))

# Add HDI data back into dataset
Area_Yield_Change_95_15 <- left_join(Area_Yield_Change_95_15,mergeset %>% dplyr::select(1,2,4,5,7))
Area_Yield_Change_95_15$HDI <- factor(Area_Yield_Change_95_15$HDI, levels = c("Low", "Medium", "High", "Very high"))
```

```{r Fig 1 Boxplots, fig.width=8, fig.height=10}
bp_area <- ggplot(data = Area_Yield_Change_95_15 %>% filter(!is.na(HDI), HDI != "N/A"), mapping = aes(x = HDI, y = log.area.dif)) +
  geom_boxplot(size = .75) + theme_classic() + geom_hline(aes(yintercept = 0)) + geom_text(aes(label = Area), size = 2, position = position_jitter(width = .3))+ labs(y = str_wrap("Log Change in per-capita Area for Staple Crop Production", width = 25))

bp_yield <- ggplot(data = Area_Yield_Change_95_15 %>% filter(!is.na(HDI), HDI != "N/A"), mapping = aes(x = HDI, y = log.yield.dif)) +
  geom_boxplot(size = .75) + theme_classic() + geom_hline(aes(yintercept = 0)) + geom_text(aes(label = Area), size = 2, position = position_jitter(width = .3)) +  labs(y = str_wrap("Log Change in Caloric Yield per hectare of Staple Crop Production", width = 25))

annotate_figure(ggarrange(bp_area,bp_yield, nrow = 2), top = "Changes to Area and Yield of Staple Crops from 1995-2015")
```


## 3 panel figure showing global trends over time: 

### A. Changes in staple crop land area, B. Global population, C. Per capita staple cropland

```{r base data inclusion, include = F}
# Components from original data sheet generation, needed for overall trends
staple_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000, Area !="Maldives", Area !="China") %>% 
  dplyr::filter(Item =="Apples"| Item == "Barley"| Item ==  "Bananas"| Item == "Cassava"| Item == "Coconuts"| 
                  Item ==  "Seed cotton"| Item ==  "Grapes"| Item ==  "Maize"| Item ==  "Millet"| Item == "Oats"| 
                  Item ==  "Onions, dry"| Item ==  "Oil palm fruit"| 
                  Item ==  "Plantains and others"| Item ==  "Potatoes"| Item ==  "Rice, paddy"| Item ==  "Sorghum"| 
                  Item == "Soybeans"| Item ==  "Sugar beet"| Item ==  "Sugar cane"| Item ==  "Sunflower seed"| 
                  Item ==  "Sweet potatoes"| Item ==  "Vegetables, fresh nes"| Item == "Wheat"| Item ==  "Yams") %>% 
  dplyr::filter(Element != "Production") 

# Renaming Year Columns
names(staple_data) = gsub(pattern = "Y", replacement = "", x = names(staple_data))
staple_data_clean <- staple_data

# Pivoting to Tall Format and selecting Area data
area.tall <- pivot_longer(data = staple_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area")%>% 
  dplyr::select(Area, Item, Year,area)
area.tall$Year <- as.numeric(area.tall$Year)


# Pivoting to Tall Format and selecting Yield data
yield.tall <- pivot_longer(data = staple_data_clean %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield") 
yield.tall$Year <- as.numeric(yield.tall$Year)

# Importing Kcal/hg data 
kcal <- read.csv("../data/kcal_staple.csv", header = T, sep = ",")

# Joining Kcal to Yield Data and converting to kcal per hectare
kcal.ha <- left_join(yield.tall,kcal %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.per.ha = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.per.ha) %>% 
  dplyr::filter(!is.na(kcal.per.ha))

# Importing Population Data and cleaning
pop_data <- read.csv("../data/fao_pop.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area !="China") %>% 
  dplyr::mutate(Population = Value*1000) %>% 
  dplyr::select(Area,Population,Year)
names(pop_data) <- as.character(names(pop_data))
pop_data_clean <- pop_data

# Combining area and yield for all staple crops, generating kcal/hectare averages for each year
area_yield_data <- left_join(area.tall,kcal.ha) %>% 
  mutate(kcal.total = area*kcal.per.ha) %>% 
  group_by(Area,Year) %>% 
  dplyr::summarise(kcal.produced.tot = sum(kcal.total,na.rm = T),
                   area.tot = sum(area,na.rm = T)) %>% 
  mutate(kcal.ha.avg = kcal.produced.tot / area.tot) %>% 
  dplyr::select(-kcal.produced.tot)

```

```{r, fig.width = 12, fig.height = 6}
area_summarized <- area.tall %>% 
  group_by(Year) %>% 
  dplyr::summarise(total.area = sum(area, na.rm = T))

pop_sum <- pop_data_clean %>% 
  group_by(Year) %>% 
  dplyr::summarise(total.pop = sum(Population, na.rm = T))

a_pop <- left_join(area_summarized,pop_sum) 
a_pop_pc <- a_pop %>% 
  dplyr::mutate(pc_area = total.area/total.pop)

a <- ggplot(data = a_pop_pc, aes(x= Year, y = total.area)) + geom_point() + theme_classic()+
              labs(y = str_wrap("Total Global Hectares for Staple Crop Production", width = 25))
b <- ggplot(data = a_pop_pc, aes(x= Year, y = total.pop)) + geom_point() + theme_classic()+
              labs(y = str_wrap("Total Global Population", width = 25))
c <- ggplot(data = a_pop_pc, aes(x= Year, y = pc_area)) + geom_point() + theme_classic()+
              labs(y = str_wrap("Global Per-capita Hectares for Staple Crop Production", width = 28))

ggarrange(a,b,c, nrow = 1)
```

```{r data verification}
# Took several global crop aggregated data for area to verify the weird dip from ~ 90-00
check_d <- read.csv("../data/fao_croparea_select.csv", header = T, sep = ",") %>% 
  dplyr::select(Area,Item,Year,Value) %>% 
  dplyr::group_by(Area,Year,Item) %>% 
  dplyr::summarize(sum_area = sum(Value, na.rm=T))
ggplot(data = check_d, aes(x= Year, y = sum_area)) + geom_point() + theme_classic() +facet_wrap(~Item, scales ="free")

```
The dip we see appears to be real, driven by primarily by rebounding cereal production combined with steady increases in other areas. Kind of beside the point for our paper, but the trends we see do appear to be real.


## 2 panel figure. 

### A. Ewer’s (1979-1999) area x yield, B current area x yield

```{r Ewers Plots,fig.width = 10}
ewersplot_79_99 <- ggplot(data = Area_Yield_Change_79_99, mapping = aes(x = log.yield.dif, y = log.area.dif)) +
  theme_classic() + geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'79 to '99 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Area_Yield_Change_79_99)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)

ewersplot_95_15 <- ggplot(data = Area_Yield_Change_95_15, mapping = aes(x = log.yield.dif, y = log.area.dif)) +
  theme_classic() + geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'95 to '15 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Area_Yield_Change_95_15)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)

ggarrange(ewersplot_79_99,ewersplot_95_15)
```


## Figure 3 split up by HDI level

```{r}
Area_Yield_Change_79_99$HDI <- factor(Area_Yield_Change_79_99$HDI, levels = c("Low", "Medium", "High", "Very high"))
Area_Yield_Change_95_15$HDI <- factor(Area_Yield_Change_95_15$HDI, levels = c("Low", "Medium", "High", "Very high"))

ewersplot_79_99_HDI <- ggplot(data = Area_Yield_Change_79_99 %>% dplyr::filter(!is.na(HDI)),
                              mapping = aes(x = log.yield.dif, y = log.area.dif)) +
  theme_classic() + geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'79 to '99 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Area_Yield_Change_79_99)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)+facet_wrap(~HDI)

ewersplot_95_15_HDI <- ggplot(data = Area_Yield_Change_95_15 %>% dplyr::filter(!is.na(HDI)),
                              mapping = aes(x = log.yield.dif, y = log.area.dif)) +
  theme_classic() + geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'95 to '15 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Area_Yield_Change_95_15)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)+facet_wrap(~HDI)

ggarrange(ewersplot_79_99_HDI,ewersplot_95_15_HDI)
```



## Figure showing impacts of nitrogen on “land sparing” relationship (compare to Fig. 3)


```{r data gen, include = F}
nitrogen_Change_95_15 = fao_staple_complete %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # Filter Countries here if wanted later
    dplyr::select(Area,Year,Population,tonnes_nitrogen) %>%    # 
    dplyr::filter(!is.na(tonnes_nitrogen)) %>%   # 
    dplyr::filter(between(Year,(1995-2),(1995+2))|between(Year,(2015-2),(2015+2))) %>%           
    dplyr::mutate(nitrogen_percap = tonnes_nitrogen/Population,
                  time = if_else(Year <= (1995+2) & Year >= (1995-2),"start.5","end.5"))%>% 
    dplyr::select(Area,time,nitrogen_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.nitrogen = mean(nitrogen_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.nitrogen)) %>% 
    dplyr::mutate(nitrogen_dif = end.5 / start.5
                  ) %>% 
    dplyr::mutate(log.nitrogen.dif = log(nitrogen_dif) 
                  ) %>% 
    dplyr::filter(!is.na(log.nitrogen.dif)) %>% 
    dplyr::select(-end.5,-start.5)

# Joining to previous data
Sparing_Nitro_95_15 <- left_join(Area_Yield_Change_95_15, nitrogen_Change_95_15)
Sparing_Nitro_95_15$HDI <- factor(Sparing_Nitro_95_15$HDI, levels = c("Low", "Medium", "High", "Very high"))
```

```{r}
ggplot(data = Sparing_Nitro_95_15, mapping = aes(x = log.yield.dif,
                                                 y = log.area.dif, 
                                                 color = log.nitrogen.dif)) + theme_classic() +
    geom_point(size = 2, alpha = .5) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + geom_text(aes(label = Area), size = 2, alpha = .7) +
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'95 to '15 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Sparing_Nitro_95_15))), color = "Log Change in per-capita nitrogen addition (tonnes)") +
    scale_color_viridis_c(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")

tidy(lm(log.area.dif ~ log.yield.dif*log.nitrogen.dif, data = Sparing_Nitro_95_15))
```
Not explicitly asked for here, but  part of our hypothesis is that this relationship between nitrogen use and land sparing differs between countries so included that here too

```{r}
ggplot(data = Sparing_Nitro_95_15 %>% filter(!is.na(HDI)), mapping = aes(x = log.yield.dif,
                                                 y = log.area.dif, 
                                                 color = log.nitrogen.dif)) + theme_classic() +
    geom_point(size = 2, alpha = .5) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + geom_text(aes(label = Area), size = 2, alpha = .7) +
    labs(x = "Log Change in Yield (kcal/ha)" , y = "Log Change in Area (ha)", 
         title = paste("'95 to '15 Changes in Yield and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Sparing_Nitro_95_15))), color = "Log Change in per-capita nitrogen addition (tonnes)") +
    scale_color_viridis_c(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")+facet_wrap(~HDI)

# Then just 2-way interactions
Anova(lm(log.area.dif ~ (log.yield.dif+log.nitrogen.dif+HDI)^2, data = Sparing_Nitro_95_15%>% filter(!is.na(HDI))))
```
Main effect of Nitrogen use on area change, interaction between yield and HDI effects on area change, and interaction between nitrogen and HDI effects on area change.


##  Figure showing impacts of exports (area x exports) split up by HDI

```{r, include = F}
Export_import_Change_95_15 = fao_staple_complete %>% 
    dplyr::select(Area,Year,Population,mean.staple.exports,mean.staple.imports) %>%    # 
    dplyr::filter(!is.na(mean.staple.exports) & !is.na(mean.staple.imports),
                  mean.staple.exports != 0, mean.staple.imports != 0) %>%   # 
    dplyr::filter(between(Year,(1995-2),(1995+2))|between(Year,(2015-2),(2015+2))) %>%           
    dplyr::mutate(export_percap = mean.staple.exports/Population, import_percap = mean.staple.imports/Population,
                  time = if_else(Year <= (1995+2) & Year >= (1995-2),"start.5","end.5"))%>% 
    dplyr::select(Area,time,export_percap,import_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.export = mean(export_percap,na.rm = T),avg.import = mean(import_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.export,avg.import))  %>% 
    dplyr::mutate(export_dif = avg.export_end.5 / avg.export_start.5,
                  import_dif = avg.import_end.5 / avg.import_start.5) %>% 
  dplyr::mutate(log.export.dif = log(export_dif), 
                  log.import.dif = log(import_dif)) %>% 
    dplyr::filter(!is.na(log.export.dif) & !is.na(log.import.dif)) %>% 
  dplyr::select(1,6,7,8,9)

Sparing_exp_imp_unrestricted_95_15 <- left_join(Area_Yield_Change_95_15,Export_import_Change_95_15)


```
```{r}
ggplot(data = Sparing_exp_imp_unrestricted_95_15 %>% filter(!is.na(HDI)), mapping = aes(x = log.export.dif,
                                                 y = log.area.dif)) + theme_classic() +
    geom_point(size = 2, alpha = .5) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + geom_text(aes(label = Area), size = 2, alpha = .7) +
    labs(y = "Log Change in Area (ha)" , 
         title = paste("'95 to '15 Changes in Export Value and Area of 23 Staple Crops", 
                       "n = ", as.character(nrow(Sparing_Nitro_95_15))), x = "Log Change in per-capita Staple Crop Export Value (USD)")+ facet_wrap(~HDI)

Anova(lm(log.area.dif ~ log.export.dif*HDI, data = Sparing_exp_imp_unrestricted_95_15%>% filter(!is.na(HDI))))
```








