---
title: "Journal Club FAO Scatterplot"
author: "Ben"
date: "1/28/2021"
output: html_document
---
## Objective
- Import and clean FAO data on yield, cropland area, and population
- Replicate/modernize Ewers (2009) scatterplot using data from 1999-20182

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = F}
library(tidyverse)
library(knitr)
library(kableExtra)
```

#### Read in Data

 23 "Staple Crops" were selected according to Ewers 2009, with yield, acreage, and population data imported for 1999-2018. 
 
 Below analyses are currently for 1999-2018

```{r Initial Data read-in, include = F}

# Updated now for all years of data
staple_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item =="Apples"| Item == "Barley"| Item ==  "Bananas"| Item == "Cassava"| Item == "Coconuts"| Item ==  "Cottonseed"| 
                  Item ==  "Grapes"| Item ==  "Maize"| Item ==  "Millet"| Item == "Oats"| Item ==  "Onions| Item ==  dry"| Item ==  "Oil palm fruit"| 
                  Item ==  "Plantains"| Item ==  "Potatoes"| Item ==  "Rice, paddy"| Item ==  "Sorghum"| 
                  Item == "Soybeans"| Item ==  "Sugar beet"| Item ==  "Sugar cane"| Item ==  "Sunflower seed"| Item ==  "Sweet potatoes"| 
                  Item ==  "Vegetables, fresh nes"| Item == "Wheat"| Item ==  "Yams") %>% 
  dplyr::filter(Element != "Production") 
names(staple_data) = gsub(pattern = "Y", replacement = "", x = names(staple_data))
str(staple_data)
pop_data <- read.csv("../data/fao_pop.csv", sep = ",", header = T) %>% 
  dplyr::mutate(Population = Value*1000) %>% 
  dplyr::select(Area,Population,Year)
names(pop_data) <- as.character(names(pop_data))

area.tall <- pivot_longer(data = staple_data %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area")
area.tall$Year <- as.numeric(area.tall$Year)

# Joining Area with population, calculating crop hectares per capita
pc.area <- left_join(area.tall,pop_data) %>% 
  dplyr::mutate(ha_percap = area/Population)%>% 
  dplyr::select(Area,Item,Year,ha_percap)

# Separating out values by year for comparison
pc.area.comparison <- pc.area %>% 
  filter(Year == 1999 | Year == 2018)

pc.area.tall <- pivot_wider(pc.area.comparison, names_from = Year, values_from = ha_percap, names_prefix = "x") %>% 
  dplyr::filter(!is.na(x2018) & !is.na(x1999) & x2018 != 0 & x1999 !=0) 

# Calculation on log(percap cropland ratio 2018: 1999) according to Ewers 2009
pc.area.dif <- pc.area.tall %>% 
  group_by(Area) %>% 
  dplyr::summarise(total.1999 = sum(x1999), total.2018 = sum(x2018)) %>% 
  dplyr::mutate(log.delta.area = log(total.2018 / total.1999)) %>% 
  dplyr::mutate(dir.area = if_else(log.delta.area >= 0, "pos", "neg")) %>% 
  dplyr::select(-total.1999,-total.2018)

```




```{r yield ratio calculation, include = F}

yield.tall <- pivot_longer(data = staple_data %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield")
yield.tall$Year <- as.numeric(yield.tall$Year)
# Manually creating calorie list because no tables exist :(
# Stolen from https://iopscience.iop.org/1748-9326/8/3/034015/media/erl472821suppdata.pdf 

kcal = data.frame(c("Apples","Barley", "Bananas","Cassava","Coconuts", "Cottonseed", "Grapes", "Maize", "Millet",
                    "Oats", "Onions, dry", "Oil palm fruit", "Plantains", "Potatoes", "Rice, paddy", "Sorghum",
                    "Soybeans", "Sugar beet", "Sugar cane", "Sunflower seed", "Sweet potatoes", "Vegetables, fresh nes",
                    "Wheat", "Yams"),  
                  c(479479.33, 3360215.93,614244.63,1058035.85,1430000.04, 4100000.00, 591589.58, 3580802.60, 3463917.52,
                    3850000.19, 394971.60, 5400000.00, 846666.67, 702122.60, 2800000.00, 3430000.40,
                    3596499.11, 700000.00, 291428.56, 2982902.68, 939797.99, 220000.00,
                    3284000.00,1090000.00))
colnames(kcal) = c("Item", "kcal.tonne")
kcal <- kcal %>% 
  dplyr::mutate(kcal.hg = kcal.tonne/10000)
# write.csv(kcal, "../data/kcal_staple.csv") # NEVER AGAIN

# Conversion
kcal.ha <- left_join(yield.tall,kcal %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.ha.cap = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.ha.cap) %>% 
  dplyr::filter(!is.na(kcal.ha.cap) )

# Pick Years
kcal.ha.1999.2018 <- kcal.ha %>% 
  filter(Year == 1999 | Year == 2018)

pc.yield.tall <- pivot_wider(kcal.ha.1999.2018, names_from = Year, values_from = kcal.ha.cap, names_prefix = "x") %>% 
  dplyr::filter(!is.na(x2018) & !is.na(x1999) & x2018 != 0 & x1999 !=0) 

kcal.yield.dif <- pc.yield.tall %>% 
  group_by(Area) %>% 
  dplyr::summarise(total.1999 = sum(x1999), total.2018 = sum(x2018)) %>% 
  dplyr::mutate(log.delta.yield = log(total.2018 / total.1999)) %>% 
  dplyr::mutate(dir.yield = if_else(log.delta.yield >= 0, "pos", "neg")) %>% 
  dplyr::select(Area,log.delta.yield,dir.yield)
```


The table below shows the values for the change in per capita cropland area and yield for the same 23 staple crops analyzed in Ewers 2009.  In this case, these "log.delta.area" represent the Y-axis of Fig. 2 from that paper, with "log.delta.yield" as the X. Notably, some of the Y values are quite extreme comparatively, with nearly twice as much cropland area per capita lost as compare to the 1979-1999 period shown in Ewers. This is in line with overall cropland area estimates that Rob showed though. "Dir" here just shows whether each value is positive or negative, though it was useful to look at overall relationships.
```{r}
combined_data <- left_join(pc.area.dif,kcal.yield.dif) 

scroll_box(
  kable(combined_data, 
        caption = "Per capita changes in cropland area and yield for 23 staple crops (1999:2018)") %>% kable_styling(),
  height = "700px",
  width = "800px")

```


In this case, the yield values are quite similar to those from Ewers.  It's possible that kcal yield per hectare for these staples hasn't changed much since 1999, but let's visualize these together in a format similar to Ewers Figure 2.
```{r}
ggplot(data = combined_data, mapping = aes(x = log.delta.yield, y = log.delta.area)) + theme_classic() +
         geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm") + xlim(-.6,1.5)+
  ylim(-3.5,1.5) + labs(x = "log(yield ratio 1999:2018)", y = "log(per capita cropland ratio 1999:2018)")
```

The models from Ewers were pretty simple, so we can look at the relationship too. 

```{r}
model <- lm(log.delta.area ~ log.delta.yield, data = combined_data)
summary(model)
```
The coefficient here is 0.122, slightly positive but not significant (p = 0.387). Again, this is only from 1999-2018, and on the 23 "staple" crops, so it doesn't mean we won't find a sparing effect later. When we break these down by a variety of other factors we may be able to gain more insight into where sparing may occur.

### Visualizing more dates
I've written a function to be able to plot these similar figures across different time points, below are a few examples!

```{r plotting function, include = F}
ewers_data <- function(staple,pop,start,end){

  
start.lab = as.name(paste("x",as.character(start),sep = ""))
end.lab = as.name(paste("x",as.character(end),sep = ""))
  #
  
  area.tall <- pivot_longer(data = staple %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area")
  area.tall$Year <- as.numeric(area.tall$Year)
  
  # Joining Area with population, calculating crop hectares per capita
  pc.area <- left_join(area.tall,pop) %>% 
    dplyr::mutate(ha_percap = area/Population)%>% 
    dplyr::select(Area,Item,Year,ha_percap)
  
  # Separating out values by year for comparison
  pc.area.comparison <- pc.area %>% 
    filter(Year == start | Year == end)
  
  pc.area.tall <- pivot_wider(pc.area.comparison, names_from = Year, values_from = ha_percap, names_prefix = "x") %>% 
    dplyr::filter(!is.na(!!start.lab) & !is.na(!!end.lab) &  !!start.lab != 0 & !!end.lab !=0) 
  
  # Calculation on log(percap cropland ratio 2018: 1999) according to Ewers 2009
  pc.area.dif <- pc.area.tall %>% 
    group_by(Area) %>% 
    dplyr::summarise(total.start = sum(!!start.lab), total.end = sum(!!end.lab)) %>%        
    dplyr::mutate(log.delta.area = log(total.end / total.start)) %>%  
    dplyr::mutate(dir = if_else(log.delta.area >= 0, "+", "-")) %>% 
    dplyr::select(-total.start,-total.end)
  
  # Yield New
  
  yield.tall <- pivot_longer(data = staple %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield")
  yield.tall$Year <- as.numeric(yield.tall$Year)
  # Manually creating calorie list because no tables exist :(
  # Stolen from https://iopscience.iop.org/1748-9326/8/3/034015/media/erl472821suppdata.pdf 
  
  kcal = data.frame(c("Apples","Barley", "Bananas","Cassava","Coconuts", "Cottonseed", "Grapes", "Maize", "Millet",
                      "Oats", "Onions, dry", "Oil palm fruit", "Plantains", "Potatoes", "Rice, paddy", "Sorghum",
                      "Soybeans", "Sugar beet", "Sugar cane", "Sunflower seed", "Sweet potatoes", "Vegetables, fresh nes",
                      "Wheat", "Yams"),  
                    c(479479.33, 3360215.93,614244.63,1058035.85,1430000.04, 4100000.00, 591589.58, 3580802.60, 3463917.52,
                      3850000.19, 394971.60, 5400000.00, 846666.67, 702122.60, 2800000.00, 3430000.40,
                      3596499.11, 700000.00, 291428.56, 2982902.68, 939797.99, 220000.00,
                      3284000.00,1090000.00))
  colnames(kcal) = c("Item", "kcal.tonne")
  kcal <- kcal %>% 
    dplyr::mutate(kcal.hg = kcal.tonne/10000)
  # write.csv(kcal, "../data/kcal_staple.csv") # NEVER AGAIN
  
  # Conversion
  kcal.ha <- left_join(yield.tall,kcal %>% dplyr::select(Item,kcal.hg)) %>% 
    dplyr::mutate(kcal.ha.cap = yield*kcal.hg) %>% 
    dplyr::select(Area, Item, Year,kcal.ha.cap) %>% 
    dplyr::filter(!is.na(kcal.ha.cap) )
  
  # Pick Years
  kcal.ha.start.end <- kcal.ha %>% 
    filter(Year == start | Year == end) 
  
  pc.yield.tall <- pivot_wider(kcal.ha.start.end, names_from = Year, values_from = kcal.ha.cap, names_prefix = "x") %>% 
    dplyr::filter(!is.na(!!start.lab) & !is.na(!!end.lab) &  !!start.lab != 0 & !!end.lab !=0) 
  
  kcal.yield.dif <- pc.yield.tall %>% 
    group_by(Area) %>% 
    dplyr::summarise(total.start = sum(!!start.lab), total.end = sum(!!end.lab)) %>% 
    dplyr::mutate(log.delta.yield = log(total.end / total.start)) %>% 
    dplyr::mutate(dir.1 = if_else(log.delta.yield >= 0, "+", "-")) %>% 
    dplyr::select(Area,log.delta.yield,dir.1)
  
  #
  
  combined_data <- left_join(pc.area.dif,kcal.yield.dif) 
  
  plot <- ggplot(data = combined_data, mapping = aes(x = log.delta.yield, y = log.delta.area)) + theme_classic() +
    geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm") + xlim(-.6,1.5)+
    ylim(-3.5,1.5) + labs(x = "log(yield ratio)", y = "log(per capita cropland ratio)", title = paste(as.character(start),"to",as.character(end)))
  
  return(plot)
}
```
```{r 1990 to 2000, include = FALSE}
# Same data but includes 1990-2018 data 
staple_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item =="Apples"| Item == "Barley"| Item ==  "Bananas"| Item == "Cassava"| Item == "Coconuts"| Item ==  "Cottonseed"| 
                  Item ==  "Grapes"| Item ==  "Maize"| Item ==  "Millet"| Item == "Oats"| Item ==  "Onions| Item ==  dry"| Item ==  "Oil palm fruit"| 
                  Item ==  "Plantains"| Item ==  "Potatoes"| Item ==  "Rice, paddy"| Item ==  "Sorghum"| 
                  Item == "Soybeans"| Item ==  "Sugar beet"| Item ==  "Sugar cane"| Item ==  "Sunflower seed"| Item ==  "Sweet potatoes"| 
                  Item ==  "Vegetables, fresh nes"| Item == "Wheat"| Item ==  "Yams") %>% 
  dplyr::filter(Element != "Production") 
names(staple_data) = gsub(pattern = "Y", replacement = "", x = names(staple_data))
staple_data_clean <- staple_data

# Load and Clean Pop data
pop_data <- read.csv("../data/fao_pop.csv", sep = ",", header = T) %>% 
  dplyr::mutate(Population = Value*1000) %>% 
  dplyr::select(Area,Population,Year)
names(pop_data) <- as.character(names(pop_data))
pop_data_clean <- pop_data

```
```{r, warning = FALSE}

# 1979 to 1999 Ewers repeat
ewers_data(staple = staple_data, pop = pop_data, start = 1976, end = 1999) 

# 2000 to 2010
ewers_data(staple = staple_data, pop = pop_data, start = 2000, end = 2010) 

# 2008 to 2018
ewers_data(staple = staple_data, pop = pop_data, start = 2008, end = 2018) 
```


Just did 10 year periods here, since the difference values change size depending on the total number of years compared (expectedly, as trends continue over time). We can easily adjust periods to fit times of interest, and compare decades or some other unit.

My guess is there are likely several agroecological factors we could look into that may explain some of the variance among countries. 
