---
title: "JClub 4/26 Updates"
author: "Ben"
date: "4/26/2021"
output:
  html_document: default
  pdf_document: default
---

```{r packages, include = F}
library(tidyverse)
library(car)
library(ggpubr)
library(kableExtra)
library(corrplot)
library(broom)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
```

```{r Function List, include = F}
# Functions: ####
  # Non-Z function 
ewers_plot_avg_noZ <- function(data, X, Y, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
 
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # FILTER COUNTRIES HERE
    dplyr::select(Area,Year,Population,!!X,!!Y) %>%    # 
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # 
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5) 
  
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif)
                  ) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  
  plot <- ggplot(data = data_cut_facet, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
    geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)
  # plot.coef <- ggplot(data = data_cut, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
  # geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm")  + labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)), "coef =", as.character(lm.coef)) )
  
  plot <- plot + 
    {if(!is.na(facet)) facet_wrap(facet)
    }
  
  return(plot)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 

ewers_plot_avg_data_noZ <- function(data, X, Y, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
  
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # FILTER COUNTRIES HERE
    dplyr::select(Area,Year,Population,!!X,!!Y) %>%    # 
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # 
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5) 
  
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif)
    ) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
 
  return(data_cut_facet)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 
# "npc" mean "non per-capita", so Znpc means the Z value is not adjusted to per-capita

    # Z npc Plotting Function
ewers_plot_avg_Znpc <- function(data, X, Y, Z, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
  z.lab = as.name(Z)
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # FILTER COUNTRIES HERE
    dplyr::select(Area,Year,Population,!!X,!!Y, !!Z) %>%    # 
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # 
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population, z_percap = !!z.lab,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap,z_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T),avg.z = mean(z_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y,avg.z))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5,
                  z_dif = avg.z_end.5 / avg.z_start.5) %>% 
    dplyr::filter(z_dif != Inf & z_dif != 0)
  
  mz <- 10*mean(data_cut$z_dif,na.rm = T)
  lmz <- mean(data_cut$z_dif,na.rm = T)/10
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif),
                  log.z.dif = ifelse(z_dif >= (mz)| z_dif <= (lmz)| is.na(z_dif), NA ,log(z_dif))) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  
  plot <- ggplot(data = data_cut_facet, mapping = aes(x = log.x.dif, y = log.y.dif, color = log.z.dif)) + theme_classic() +
    geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut))), color = z.lab) +
    scale_color_viridis_c(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")
  # plot.coef <- ggplot(data = data_cut, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
  # geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm")  + labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)), "coef =", as.character(lm.coef)) )
  
  plot <- plot + 
    {if(!is.na(facet)) facet_wrap(facet)
    }
  
  return(plot)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 

    # Z npc Data Function
ewers_plot_avg_data_Znpc <- function(data, X, Y, Z, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
  z.lab = as.name(Z)
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    dplyr::select(Area,Year,Population,!!X,!!Y, !!Z) %>%    # grab only relevant data
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # filters out missing values
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population, z_percap = !!z.lab,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap,z_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T),avg.z = mean(z_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y,avg.z))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5,
                  z_dif = avg.z_end.5 / avg.z_start.5)%>% 
    dplyr::filter(z_dif != Inf & z_dif != 0)
  
  mz <- 10*mean(data_cut$z_dif,na.rm = T)
  lmz <- mean(data_cut$z_dif,na.rm = T)/10
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif),
                  log.z.dif = ifelse(z_dif >= (mz)| z_dif <= (lmz)| is.na(z_dif), NA ,log(z_dif))) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  data_clean <- data_cut_facet %>% 
    dplyr::select(Area,log.x.dif,log.y.dif,log.z.dif,HDI,Developed,Continent)
  
  return(data_clean)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 

    # Z pc Plotting Function
ewers_plot_avg_Zpc <- function(data, X, Y, Z, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
  z.lab = as.name(Z)
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # FILTER COUNTRIES HERE
    dplyr::select(Area,Year,Population,!!X,!!Y, !!Z) %>%    # 
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # 
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population, z_percap = !!z.lab/Population,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap,z_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T),avg.z = mean(z_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y,avg.z))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5,
                  z_dif = avg.z_end.5 / avg.z_start.5) %>% 
    dplyr::filter(z_dif != Inf & z_dif != 0)
  
  mz <- 10*mean(data_cut$z_dif,na.rm = T)
  lmz <- mean(data_cut$z_dif,na.rm = T)/10
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif),
                  log.z.dif = ifelse(z_dif >= (mz)| z_dif <= (lmz)| is.na(z_dif), NA ,log(z_dif))) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  
  plot <- ggplot(data = data_cut_facet, mapping = aes(x = log.x.dif, y = log.y.dif, color = log.z.dif)) + theme_classic() +
    geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + 
    labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut))), color = z.lab) +
    scale_color_viridis_c(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")
  # plot.coef <- ggplot(data = data_cut, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
  # geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm")  + labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)), "coef =", as.character(lm.coef)) )
  
  plot <- plot + 
    {if(!is.na(facet)) facet_wrap(facet)
    }
  
  return(plot)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 

    # Z pc Data Function
ewers_plot_avg_data_Zpc <- function(data, X, Y, Z, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
  z.lab = as.name(Z)
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    dplyr::select(Area,Year,Population,!!X,!!Y, !!Z) %>%    # grab only relevant data
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # filters out missing values
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population, z_percap = !!z.lab/Population,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap,z_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T),avg.z = mean(z_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y,avg.z))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5,
                  z_dif = avg.z_end.5 / avg.z_start.5)%>% 
    dplyr::filter(z_dif != Inf & z_dif != 0)
  
  mz <- 10*mean(data_cut$z_dif,na.rm = T)
  lmz <- mean(data_cut$z_dif,na.rm = T)/10
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif),
                  log.z.dif = ifelse(z_dif >= (mz)| z_dif <= (lmz)| is.na(z_dif), NA ,log(z_dif))) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(2,16,18,19) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  data_clean <- data_cut_facet %>% 
    dplyr::select(Area,log.x.dif,log.y.dif,log.z.dif,HDI,Developed,Continent)
  
  return(data_clean)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 

ewers_plot_avg_commod <- function(data, X, Y, start, end, facet= NA){
  
  start.lab = as.name(paste("x",as.character(start),sep = ""))
  end.lab = as.name(paste("x",as.character(end),sep = ""))
  x.lab = as.name(X)
  y.lab = as.name(Y)
 
  #
  # Cut data for years and variables only, adjust for per capita
  data_cut = data %>% 
    # dplyr::filter(export_value_usd >= 200000) %>%  # FILTER COUNTRIES HERE
    dplyr::select(Area,Year,Population,!!X,!!Y) %>%    # 
    dplyr::filter(!is.na(!!x.lab) & !is.na(!!y.lab)) %>%   # 
    dplyr::filter(between(Year,(start-2),(start+2))|between(Year,(end-2),(end+2)))   %>%                            # Select Years
    dplyr::mutate(x_percap = !!x.lab/Population, y_percap = !!y.lab/Population,
                  time = if_else(Year <= (start+2) & Year >= (start-2),"start.5","end.5"))%>% #Note that X here now is not per-capita
    dplyr::select(Area,time,x_percap,y_percap) %>% 
    dplyr::group_by(Area,time) %>% 
    dplyr::summarize(avg.x = mean(x_percap,na.rm = T),avg.y = mean(y_percap,na.rm = T))%>% 
    pivot_wider(names_from = time, values_from = c(avg.x,avg.y))  %>% 
    dplyr::mutate(x_dif = avg.x_end.5 / avg.x_start.5,
                  y_dif = avg.y_end.5 / avg.y_start.5) 
  
  data_cut = data_cut %>% 
    dplyr::mutate(log.x.dif = log(x_dif), 
                  log.y.dif = log(y_dif)
                  ) %>% 
    dplyr::filter(!is.na(log.x.dif) & !is.na(log.y.dif))
  
  # Fix how model runs later
  #lm = lm(log.y.dif ~ log.x.dif, data = data_cut)
  
  #lm.coef = round(as.data.frame(coef(lm))[2,1],digits = 3)
  head(data)
  merge_set <- data %>% 
    dplyr::select(1,7,8,9) %>% 
    distinct(.)
  head(merge_set)
  data_cut_facet <- left_join(data_cut,merge_set, by = "Area") %>% 
    dplyr::filter(!is.na(HDI) & HDI != "N/A")
  
  
  plot <- ggplot(data = data_cut_facet, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
    geom_point(size = 3, alpha = .3) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
    geom_smooth(method = "lm")  + geom_text(aes(label = Area), size = 2, alpha = .7) +
    labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)))) + 
    theme(legend.position = "bottom")+ ylim(-2.5,1) + xlim(-1.5,1)
  # plot.coef <- ggplot(data = data_cut, mapping = aes(x = log.x.dif, y = log.y.dif)) + theme_classic() +
  # geom_point() + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + geom_smooth(method = "lm")  + labs(x = x.lab, y = y.lab, title = paste(as.character(start),"to",as.character(end), "n = ", as.character(nrow(data_cut)), "coef =", as.character(lm.coef)) )
  
  plot <- plot + 
    {if(!is.na(facet)) facet_wrap(facet)
    }
  
  return(plot)
  # if_else(is.na(lm.coef), return(plot), return(plot.coef))
} 
```

## JC Project 4/26
Ben's list of things to do for Journal club prior to meeting on 4/29

1. Generate new datasets for crop subsets
2. Generate full corrplot for relevant responses
3. Summarize Models for 1995-2015 on interactive effects of Nitrogen, Exp/Imp, and Pesticide use on area changes
4. Re-examine Z-level Factors non-transformed
 

### 1. Generate new datasets for crop subsets

#### Goal: Have datasets available where we can examine effects on only a subset of staple crops, i.e. those primarily GM in certain countries
From what I've seen, the main 4 GM crops grown globally are soybean, cotton, maize, and canola, and since canola isn't one of those staple crops I'm going to create a few datasets looking at the other 3 together and individuallly.

Big disclaimer here is that "cotton" is actually "Seed cotton" aka calories from cottonseed oil, so that may be unrepresentative when it comes to actual cotton production or yield calculations.

```{r GM data, include = F}
#OG Dataset
fao_updated_staple = read.csv("../data/fao_updated_staple.csv", header = T, sep = ",")
fao_updated_staple$HDI  <- factor(fao_updated_staple$HDI, levels = c("Low", "Medium", "High", "Very high"))
merge_set <- read.csv("../data/mergeset.csv", sep = ",", header = T)

#kcal and Pop data for reference
kcal_data <- read.csv("../data/kcal_staple.csv", sep = ",", header = T)
pop_data <- read.csv("../data/fao_pop.csv", sep = ",", header = T) %>% 
  dplyr::mutate(Population = Value*1000) %>% 
  dplyr::select(Area,Population,Year)
names(pop_data) <- as.character(names(pop_data))

# soybeans
soy_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item == "Soybeans") %>% 
  dplyr::filter(Element != "Production") 
names(soy_data) = gsub(pattern = "Y", replacement = "", x = names(soy_data))
soy_data_clean <- soy_data
soy_area_tall <- pivot_longer(data = soy_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area.harv")%>%
  dplyr::select(Area, Item, Year,area.harv)
soy_area_tall$Year <- as.numeric(soy_area_tall$Year)
head(soy_area_tall)

soy_yield_tall <- pivot_longer(data = soy_data_clean %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield") 
soy_yield_tall$Year <- as.numeric(soy_yield_tall$Year)
soy_yield_full <- left_join(soy_yield_tall,kcal_data %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.ha.avg = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.ha.avg) %>% 
  dplyr::filter(!is.na(kcal.ha.avg))

soy_full <- left_join(soy_area_tall,soy_yield_full) %>% 
  left_join(.,pop_data) %>% 
  left_join(.,merge_set)
write.csv(soy_full,"../data/soy_data.csv")

# maize
maize_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item == "Maize") %>% 
  dplyr::filter(Element != "Production") 
names(maize_data) = gsub(pattern = "Y", replacement = "", x = names(maize_data))
maize_data_clean <- maize_data
maize_area_tall <- pivot_longer(data = maize_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area.harv")%>%
  dplyr::select(Area, Item, Year,area.harv)
maize_area_tall$Year <- as.numeric(maize_area_tall$Year)
head(maize_area_tall)

maize_yield_tall <- pivot_longer(data = maize_data_clean %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield") 
maize_yield_tall$Year <- as.numeric(maize_yield_tall$Year)
maize_yield_full <- left_join(maize_yield_tall,kcal_data %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.ha.avg = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.ha.avg) %>% 
  dplyr::filter(!is.na(kcal.ha.avg))

maize_full <- left_join(maize_area_tall,maize_yield_full) %>% 
  left_join(.,pop_data)%>% 
  left_join(.,merge_set)
write.csv(maize_full,"../data/maize_data.csv")

# cotton
cotton_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item == "Seed cotton") %>% 
  dplyr::filter(Element != "Production") 
names(cotton_data) = gsub(pattern = "Y", replacement = "", x = names(cotton_data))
cotton_data_clean <- cotton_data
cotton_area_tall <- pivot_longer(data = cotton_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area.harv")%>%
  dplyr::select(Area, Item, Year,area.harv)
cotton_area_tall$Year <- as.numeric(cotton_area_tall$Year)
head(cotton_area_tall)

cotton_yield_tall <- pivot_longer(data = cotton_data_clean %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield") 
cotton_yield_tall$Year <- as.numeric(cotton_yield_tall$Year)
cotton_yield_full <- left_join(cotton_yield_tall,kcal_data %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.ha.avg = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.ha.avg) %>% 
  dplyr::filter(!is.na(kcal.ha.avg))

cotton_full <- left_join(cotton_area_tall,cotton_yield_full) %>% 
  left_join(.,pop_data)%>% 
  left_join(.,merge_set)
write.csv(cotton_full,"../data/cotton_data.csv")

# GM3 (all three)
GM3_data <- read.csv("../data/Production_Crops_E_All_Data_NOFLAG.csv", sep = ",", header = T) %>% 
  dplyr::filter(Area.Code < 1000) %>% 
  dplyr::filter(Item == "Seed cotton"|Item =="Soybeans"|Item == "Maize") %>% 
  dplyr::filter(Element != "Production") 
names(GM3_data) = gsub(pattern = "Y", replacement = "", x = names(GM3_data))
GM3_data_clean <- GM3_data
GM3_area_tall <- pivot_longer(data = GM3_data_clean %>% filter(Element == "Area harvested"),cols = 8:66, names_to = "Year", values_to = "area.harv")%>%
  dplyr::select(Area, Item, Year,area.harv)
GM3_area_tall$Year <- as.numeric(GM3_area_tall$Year)
head(GM3_area_tall)

GM3_yield_tall <- pivot_longer(data = GM3_data_clean %>% filter(Element == "Yield"),cols = 8:66, names_to = "Year", values_to = "yield") 
GM3_yield_tall$Year <- as.numeric(GM3_yield_tall$Year)
GM3_yield_full <- left_join(GM3_yield_tall,kcal_data %>% dplyr::select(Item,kcal.hg)) %>% 
  dplyr::mutate(kcal.per.ha = yield*kcal.hg) %>% 
  dplyr::select(Area, Item, Year,kcal.per.ha) %>% 
  dplyr::filter(!is.na(kcal.per.ha))

GM3_full <- left_join(GM3_area_tall,GM3_yield_full) %>% 
  left_join(.,pop_data)%>% 
  left_join(.,merge_set)

GM3_full_summed <- GM3_full %>% 
  mutate(kcal.total = area.harv*kcal.per.ha) %>% 
  group_by(Area,Year) %>% 
  dplyr::summarise(kcal.produced.tot = sum(kcal.total,na.rm = T),
                   area.tot = sum(area.harv,na.rm = T)) %>% 
  mutate(kcal.ha.avg = kcal.produced.tot / area.tot) %>% 
  dplyr::select(-kcal.produced.tot)

write.csv(GM3_full,"../data/GM3_data.csv")
```

Not shown here, but have datasets now available that have each of the three individually and combined for analysis. 

Quick example, lets look at the 1979-1999 and 1995-2015 Ewers relationships if it was ONLY SOYBEANS!
(made another new little function, only other change is "area.tot" is now "area.harv")

#### Just Soy 

```{r}
soy1 <- ewers_plot_avg_commod(data = soy_full %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                   X = "kcal.ha.avg",
                   Y = "area.harv",
                   start = 1979,
                   end = 1999) 
soy2 <- ewers_plot_avg_commod(data = soy_full %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                   X = "kcal.ha.avg",
                   Y = "area.harv",
                   start = 1995,
                   end = 2015)
ggarrange(soy1,soy2, common.legend = T)
```

Interesting but unsurprising that individual commodities dont follow global trends!

## 2. Generate full corrplot for relevant responses
Looking first at untransformed values here.

#### 1995

```{r 1995}
corrplot(cor(drop_na(fao_updated_staple %>% filter(Year == 1995) %>% 
               dplyr::select(3,area.tot,tonnes_nitrogen,CO2_eq_emissions,tonnes_pest_total,kcal.ha.avg,mean.staple.exports,mean.staple.imports))))

```


#### 2015

```{r}
corrplot(cor(drop_na(fao_updated_staple %>% filter(Year == 2015) %>% 
               dplyr::select(3,tonnes_nitrogen,CO2_eq_emissions,tonnes_pest_total,kcal.ha.avg,mean.staple.exports,mean.staple.imports))))
```



#### 1995 per capita

Below two plots are raw values / Population 

```{r}
corrplot(cor(drop_na(fao_updated_staple %>% filter(Year == 1995) %>% 
               dplyr::select(3,tonnes_nitrogen,CO2_eq_emissions,tonnes_pest_total,kcal.ha.avg,mean.staple.exports,mean.staple.imports) %>% 
              dplyr::mutate(tonnes_nitrogen.pc = tonnes_nitrogen/Population,
                            CO2_eq_emissions.pc = CO2_eq_emissions/Population,
                            tonnes_pest_total.pc = tonnes_pest_total/Population,
                            kcal.ha.avg1 = kcal.ha.avg,
                            mean.staple.exports.pc = mean.staple.exports/Population,
                            mean.staple.imports.pc = mean.staple.imports/Population) %>% 
                dplyr::select(-2:-7))))
```

#### 2015 per capita

```{r}
corrplot(cor(drop_na(fao_updated_staple %>% filter(Year == 2015) %>% 
               dplyr::select(3,tonnes_nitrogen,CO2_eq_emissions,tonnes_pest_total,kcal.ha.avg,mean.staple.exports,mean.staple.imports) %>% 
              dplyr::mutate(tonnes_nitrogen.pc = tonnes_nitrogen/Population,
                            CO2_eq_emissions.pc = CO2_eq_emissions/Population,
                            tonnes_pest_total.pc = tonnes_pest_total/Population,
                            kcal.ha.avg1 = kcal.ha.avg,
                            mean.staple.exports.pc = mean.staple.exports/Population,
                            mean.staple.imports.pc = mean.staple.imports/Population) %>% 
                dplyr::select(-2:-7))))
```

## 3. Summarize All Existing Models for 1995-2015 on interactive effects of Nitrogen, Exp/Imp, and Pesticide use on area changes

### Figure 1: Per capita change in area used to grow and yield of 23 staple crops

First Figure is the basic comparison of Ewers 2009 style analysis with updated 1995-2015 data. 
Reminder, our method now includes averaging of two years before and after stated time (i.e. "1995" is the average of 1993-1997)
```{r, echo = F}
base_ewers_79_99 <- ewers_plot_avg_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                   X = "kcal.ha.avg",
                   Y = "area.tot",
                   start = 1979,
                   end = 1999) 
base_ewers_95_15 <- ewers_plot_avg_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                       X = "kcal.ha.avg",
                                       Y = "area.tot",
                                       start = 1995,
                                       end = 2015) 
JCfig_1 <- ggarrange(base_ewers_79_99,base_ewers_95_15, common.legend = T)
JCfig_1
```

These models represent effects of "kcal.ha.avg" (yield metric) on area used to grow staple crops
#### Model Summaries

```{r, include = F}
ewers_79_99_data <- ewers_plot_avg_data_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                        X = "kcal.ha.avg",
                                        Y = "area.tot",
                                        start = 1979,
                                        end = 1999)

ewers_95_15_data <- ewers_plot_avg_data_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                            X = "kcal.ha.avg",
                                            Y = "area.tot",
                                            start = 1995,
                                            end = 2015)
```


```{r Figure 1 models}

    # Old Ewers Model (Area Change ~ Yield Change 1979-1999)
Anova(lm(log.y.dif ~ log.x.dif, data = ewers_79_99_data))
    # New Ewers Model (Area Change ~ Yield Change 1995-2015)
Anova(lm(log.y.dif ~ log.x.dif, data = ewers_95_15_data))
```

We find a slightly negative but non-significant relationship from the 1979-1999 period between yield and total area, but a significant positive relationship from 1995-2015.

As a reminder, Ewers found a slight negative relationship in their analysis, so we will need to address this later given our difference there. I think our year averaging technique is superior, but may need to incorporate what they did for the Soviet Union (we have like 40 more countries than them too).

#### HDI Faceting same Values

```{r Ewers style faceted, echo = F}
facet_ewers_79_99 <- ewers_plot_avg_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                       X = "kcal.ha.avg",
                                       Y = "area.tot",
                                       facet = "HDI",
                                       start = 1979,
                                       end = 1999) 
facet_ewers_95_15 <- ewers_plot_avg_noZ(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                       X = "kcal.ha.avg",
                                       Y = "area.tot",
                                       facet = "HDI",
                                       start = 1995,
                                       end = 2015) 
JCfig_1_opt2 <- ggarrange(facet_ewers_79_99, facet_ewers_95_15, common.legend = T)
JCfig_1_opt2
```

This shows that positive relation is driven primarily by High and Very high HDI contries, and could help justify our investigation of countries by HDI and other explaining factors.


### Nitrogen

First Figure is a recreation of the Ewers style plots, with change in nitrogen use on the Z. Model summaries below that look at the interaction between changes in yield and changes in nitrogen use on area used to grow staple crops.


```{r Nitrogen Data gen and plots, echo = F}
data79_99_nitro <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                         X = "kcal.ha.avg",
                                         Y = "area.tot",
                                         Z = "tonnes_nitrogen",
                                         start = 1979,
                                         end = 1999)

data95_15_nitro <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                         X = "kcal.ha.avg",
                                         Y = "area.tot",
                                         Z = "tonnes_nitrogen",
                                         start = 1995,
                                         end = 2015)
# Ewers Style Interaction Models
early_nitro <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                    X = "kcal.ha.avg",
                                    Y = "area.tot",
                                    Z = "tonnes_nitrogen",
                                    start = 1979,
                                    end = 1999)

recent_nitro <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                     X = "kcal.ha.avg",
                                     Y = "area.tot",
                                     Z = "tonnes_nitrogen",
                                     start = 1995,
                                     end = 2015)

ggarrange(early_nitro,recent_nitro, nrow =1, common.legend = T)

```
 
```{r}
# 1979-1999 Area ~ Yield * Nitrogen
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data79_99_nitro))

# 1995-2015 Area ~ Yield * Nitrogen
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data95_15_nitro))
```

### Area Changes ~ Nitrogen Use 1995-2015

```{r, echo = F}
nitro_on_area_95_15 <- ggplot(data = data95_15_nitro, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Nitrogen Applied in Agriculture", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")
nitro_on_area_95_15
Anova(lm(log.y.dif ~ log.z.dif, data = data95_15_nitro))
```

#### Same again Split by HDI this time

```{r, echo = F}
nitro_on_area_95_15_HDI <- ggplot(data = data95_15_nitro, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + 
  labs(x = "Change in Nitrogen Applied in Agriculture", y = "Change in Area for Staple Crop Production") + geom_text(aes(label = Area), size = 3, alpha = .3)+
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom") +
  facet_wrap(~HDI)
nitro_on_area_95_15_HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = data95_15_nitro))
```

### Imports and Exports 

#### Exports

```{r Exp Data and First Ewers, echo = F}
data79_99_exp <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                     X = "kcal.ha.avg",
                                     Y = "area.tot",
                                     Z = "mean.staple.exports",
                                     start = 1979,
                                     end = 1999)

data95_15_exp <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                     X = "kcal.ha.avg",
                                     Y = "area.tot",
                                     Z = "mean.staple.exports",
                                     start = 1995,
                                     end = 2015)

# Plots and models
early_exports <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.exports",
               start = 1979,
               end = 1999)

recent_exports <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.exports",
               start = 1995,
               end = 2015)

ggarrange(early_exports,recent_exports, nrow =1, common.legend = T)
```

^^^ Lots of missing values above

Analysis

```{r Exp Analyses}
# 1979-1999 Area ~ Yield * Exports
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data79_99_exp))

# 1995-2015 Area ~ Yield * Exports
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data95_15_exp))
```

#### Imports

```{r, echo = F}
data79_99_imp <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                     X = "kcal.ha.avg",
                                     Y = "area.tot",
                                     Z = "mean.staple.imports",
                                     start = 1979,
                                     end = 1999)

data95_15_imp <- ewers_plot_avg_data_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
                                     X = "kcal.ha.avg",
                                     Y = "area.tot",
                                     Z = "mean.staple.imports",
                                     start = 1995,
                                     end = 2015)

early_imp <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.imports",
               start = 1979,
               end = 1999)


recent_imp <- ewers_plot_avg_Zpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.imports",
               start = 1995,
               end = 2015)

ggarrange(early_imp,recent_imp, nrow =1, common.legend = T)
```

```{r Imp analyses}

# 1979-1999 Area ~ Yield * Imports
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data79_99_imp))


# 1995-2015 Area ~ Yield * Imports
tidy(lm(log.y.dif ~ log.x.dif*log.z.dif, data = data95_15_imp))
```

#### Area Changes ~ Exports/Imports 1995-2015

```{r solo Exp/Imp Plots, echo = F}
exp_on_area_95_15 <- ggplot(data = data95_15_exp, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Value of Staple Crop Exports", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")

imp_on_area_95_15 <- ggplot(data = data95_15_imp, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Value of Staple Crop imports", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")

ggarrange(exp_on_area_95_15,imp_on_area_95_15,nrow = 1)
```

```{r Solo Exp Imp analyses}
# 1995-2015 Area Change ~ Change in Exports
tidy(lm(log.y.dif ~ log.z.dif, data = data95_15_exp))

# 1995-2015 Area Change ~ Change in Imports
tidy(lm(log.y.dif ~ log.z.dif, data = data95_15_imp))

```

#### Export/Import Split by HDI

```{r, echo = F}
exp_on_area_95_15_HDI <- ggplot(data = data95_15_exp, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + 
  labs(x = "Change in Value of Staple Crop Exports", y = "Change in Area for Staple Crop Production") +geom_text(aes(label = Area), size = 3, alpha = .3)+
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom") +
  facet_wrap(~HDI)

imp_on_area_95_15_HDI <- ggplot(data = data95_15_imp, mapping = aes(x = log.z.dif, y = log.y.dif)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(method = "lm", alpha = .2)  + 
  labs(x = "Change in Value of Staple Crop imports", y = "Change in Area for Staple Crop Production") +geom_text(aes(label = Area), size = 3, alpha = .3)+
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom") +
  facet_wrap(~HDI)

ggarrange(exp_on_area_95_15_HDI,imp_on_area_95_15_HDI, nrow = 1)
```
Analyses

```{r}
# 1995-2015 Area Change ~ Change in Exports * HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = data95_15_exp))

# 1995-2015 Area Change ~ Change in Imports * HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = data95_15_imp))
```

## 4. Examine Z factors non-transformed

This is now looking at some plots and figures without the Z level responses (i.e. nitrogen) scaled to per-capita. These still represent log(late/early) values to reflect change over time, they values are just not adjusted to population for each region. This may be a bit useless in the long run, but might help us identify some of the biggest changes in practices.

#### 
```{r, echo = F}
unscaled_nitro <- ewers_plot_avg_data_Znpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "tonnes_nitrogen",
               start = 1995,
               end = 2015)
unscaled_exports <- ewers_plot_avg_data_Znpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.exports",
               start = 1995,
               end = 2015)
unscaled_imports <- ewers_plot_avg_data_Znpc(data = fao_updated_staple %>% filter(Area != "Brunei Darussalam", Area != "Maldives"), 
               X = "kcal.ha.avg",
               Y = "area.tot",
               Z = "mean.staple.imports",
               start = 1995,
               end = 2015)

ggplot(data = unscaled_nitro, mapping = aes(x = log.z.dif, y = log.y.dif, color = HDI)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(aes(color = NA),method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Nitrogen Additions (unscaled)", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")


ggplot(data = unscaled_exports, mapping = aes(x = log.z.dif, y = log.y.dif, color = HDI)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(aes(color = NA),method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Value of Staple Crop Exports (unscaled)", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")

ggplot(data = unscaled_imports, mapping = aes(x = log.z.dif, y = log.y.dif, color = HDI)) + theme_classic() +
  geom_point(size = 3, alpha = .75) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_smooth(aes(color = NA),method = "lm", alpha = .2)  + geom_text(aes(label = Area), size = 3, alpha = .3)+
  labs(x = "Change in Value of Staple Crop Imports(unscaled)", y = "Change in Area for Staple Crop Production") +
  scale_color_viridis_d(begin = 0.2,end = .9,option = "inferno", na.value = "grey50") + theme(legend.position = "bottom")
```
```{r}
# 1995-2015 Area Change ~ Change in Nitrogen Applied (tonnes unscaled) * HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = unscaled_nitro))

# 1995-2015 Area Change ~ Change in Exports (unscaled) * HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = unscaled_exports))

# 1995-2015 Area Change ~ Change in Imports (unscaled) * HDI
Anova(lm(log.y.dif ~ log.z.dif*HDI, data = unscaled_imports))
```

At first glance there doesn't really seem to be that much of a difference with or without scaling (given the ratio is log transformed) so we may want to look into another way of investigating this. Instead of the ration (late/early) we could simply look at one time points completely raw or log-transformed values for correlations with change in area (i.e. only the countries with the most tonnes of nitrogen added had changes to area).

